---
title: "R Notebook"
output: html_notebook
---


```{r message=FALSE}
library(covid19.analytics)
library(tidyverse)
library(lubridate)
library(magrittr)
```

```{r}
# read daily data into an object named covid_daily

covid_daily <- covid19.US.data() 


# read the cumulative state level data

covid_cum <- covid19.JHU.data() %>% 
  janitor::clean_names() %>% 
  as_tibble()

# create a data vector for use in reports

last_update <-
  covid_cum$Last_Update[1] %>% 
  as_date() %>% 
    format("%A, %B %d")
```

```{r}
# This function will make it easier to "not choose" records

`%notin%` <- Negate(`%in%`)
```

# Data first glance

```{r}
covid_daily %>% 
  distinct(Country_Region)
  # count(Country_Region)

unique(covid_daily$Country_Region)

covid_daily %>% 
  filter(Province_State %in% c(state.name, "District of Columbia")) %>% 
  select(-Country_Region)
         
c(state.name, "District of Columbia")
```

Let's choose observations from the 50 states and Washington DC. Base R includes a vector called state.name in datasets.  However, state does not include DC.  Let's also exclude Country_Region.

```{r}
covid_daily_us <- 
covid_daily %>% 
  filter(Province_State %in% c(state.name, "District of Columbia")) %>% 
  select(-Country_Region)
```

Confirm 50 states and DC

```{r}
covid_daily_us %>%
  distinct(Province_State)

```



```{r}
covid_cum %>% 
  filter(Country_Region == "US") %>% 
  distinct(Province_State)

covid_cum
```

Note that the covid_daily data don't have county name column, but the covid_sum does.  We will need to join them to get the county names.  What are the keys?

## Are the data tidy?

No, so now what

```{r}
covid_daily_us_long <- 
covid_daily_us %>% 
  pivot_longer(cols = -c(Province_State, Lat, Long_),
               names_to = "date",
               values_to = "value")
```
Don't have county names.
Get the county name from covid_sum.  We need only Admin2, Long_, and Lat from covid_cum 

```{r}
covid_daily_us_long %<>% 
  # pivot_longer(cols = -c(Province_State, Lat, Long_),
  #              names_to = "date",
  #              values_to = "value") %>% 
  right_join(covid_cum %>% 
               select(Admin2, Lat, Long_),
            by = c("Lat", "Long_")) %>% 
  # remove Lat Long_ 
  select(-Lat, - Long_) %>% 
  rename(county = Admin2, state = Province_State) %>% 
  #relocate the county column
  relocate(county, .after = state)
```

Let's look at Prince Edward county data

```{r}
covid_daily_us_long %>% 
  filter(state == "Virginia", county == "Buckingham")
```

Note the number of observations is double the number of days. Why do you think that is?

There are two observations for each month one for confirmed on for deaths.  So we need to create a "long" character vector...

How can we arrange the data to simplify the creation of this 

```{r}
covid_daily_us_long %>% 
  arrange(state, county, date)
```

Should have even number of values...

```{r}
covid_daily_us_long %>% 
  arrange(state, county, date) %>% 
  mutate(type = rep(c("confirmed", "deaths"),
                    length.out = length(county)))
```

Now we need to pivot_wide---

```{r}
# covid_daily_us_wide <- 
covid_daily_us_long %>% 
  arrange(state, county, date) %>% 
  mutate(type = rep(c("confirmed", "deaths"),
                    length.out = length(county))) %>% 
  pivot_wider(names_from = type,
              values_from = value)
```

Find counties with issues

```{r}
covid_daily_us_long %>% 
  drop_na() %>% 
  arrange(state, county, date) %>% 
  mutate(type = rep(c("confirmed", "deaths"),
                    length.out = length(county))) %>% 
  pivot_wider(names_from = type,
              values_from = value,
              values_fn = length) %>% 
  filter(confirmed > 1) %>% 
  distinct(county) %>% 
  pull() %>% 
  as.character()-> doubles
```

remove problematic counties

```{r}
covid_daily_us_long %>% 
  filter(county %notin% doubles) %>% 
  drop_na() %>% 
  arrange(state, county, date) %>% 
  mutate(type = rep(c("confirmed", "deaths"),
                    length.out = length(county))) %>% 
  pivot_wider(names_from = type,
              values_from = value,
              values_fn = length) %>% 
  filter(deaths > 1 | confirmed > 1)
```

0 rows means problem counties have been removed.

```{r}
covid_daily_us_final <- 
covid_daily_us_long %>% 
  filter(county %notin% doubles) %>% 
  drop_na() %>% 
  arrange(state, county, date) %>% 
  mutate(type = rep(c("confirmed", "deaths"),
                    length.out = length(county))) %>% 
  pivot_wider(names_from = type,
              values_from = value)
```


```{r}
covid_daily_us_final %>% 
  mutate(stay_home = ifelse(date >= "2020-03-16" & date <= "2020-06-15", 1, 0)) %>% 
  filter(date >= "2020-06-15")
```



```{r}
covid_daily_us_final %>% 
  filter(state %in% c("New York", "California", "Florida", "South Carolina")) %>% 
  filter(date >= "2020-03-01", date <= "2020-06-30",
        confirmed >= 1) %>%
  group_by(state, date) %>% 
  mutate(confirmed = sum(confirmed),
         deaths = sum(deaths)) %>% 
  ggplot(aes(x = as.Date(date), y = deaths, color = state)) + 
  geom_line() #+
  # geom_line(aes(y = confirmed, x = as.Date(date), color = state)) +
  # scale_y_log10() +
```

```{r}
covid_daily_us_final %>% 
  group_by(county) %>% 
  mutate(chg_deaths = deaths - lag(deaths)) %>% 
  ungroup() %>% 
  group_by(date) %>% 
  mutate(tot_deaths = sum(deaths),
         new_deaths = sum(chg_deaths)) %>% 
  ggplot(aes(x = as.Date(date), y = new_deaths)) + 
  geom_line()
```


Created time series data by county.
```{r}
# jhu %>% 
#   select(Admin2, Lat, Long_) %>% 
#   left_join(covid, by = c("Lat", "Long_")) %>% 
#   # filter for states and DC
#   filter(Province_State %in% c(state.name, "District of Columbia")) %>% 
#   select(-c(Lat, Long_, Country_Region)) %>% 
#   pivot_longer(cols = -c(Admin2, Province_State),
#                names_to = "date",
#                values_to = "value") %>% 
#   # group_by(Admin2, date) %>% 
#   # pivot_wider(names_from = Admin2,
#   #             values_from = value) %>% 
#   janitor::clean_names() %>% 
#   arrange(province_state, admin2, date) %>% 
#   mutate(type = rep(c("confirmed", "deaths"), length.out = length(admin2))) %>% 
#   filter(admin2 %notin% doubles) %>% 
#   # filter(value > 10)
#   # distinct() %>% 
#   pivot_wider(names_from = type,
#               values_from = value) 
  
```

Attempt to Remove double counted counties

```{r}
# jhu %>% 
#   filter(Admin2 == "Dukes and Nantucket")
```





```{r}
# df %>% 
#   filter(confirmed > 1) %>% 
#   distinct(admin2) %>% 
#   pull(admin2) %>% 
#   as.character() -> doubles

# doubles combine counties 

# remove those admin2 from data.
```

find duplicate rows

```{r}
# jhu %>% 
#   select(Admin2, Lat, Long_) %>% 
#   left_join(covid, by = c("Lat", "Long_")) %>% 
#   # filter for states and DC
#   filter(Province_State %in% c(state.name, "District of Columbia")) %>% 
#   select(-c(Lat, Long_, Country_Region)) %>% 
#   filter(Province_State == "Utah") %>% 
#   pivot_longer(-c(Province_State, Admin2),
#               names_to = "date",
#               values_to = "value")
```



```{r}
# jhu %>% 
#   select(Admin2, Lat, Long_) %>% 
#   left_join(covid, by = c("Lat", "Long_")) %>% 
#   # filter for states and DC
#   filter(Province_State %in% c(state.name, "District of Columbia")) %>% 
#   select(-c(Lat, Long_, Country_Region)) %>% 
#   pivot_longer(cols = -c(Admin2, Province_State),
#                names_to = "date",
#                values_to = "value") %>% 
#   # group_by(Admin2, date) %>% 
#   # pivot_wider(names_from = Admin2,
#   #             values_from = value) %>% 
#   janitor::clean_names()
```

```{r}
# jhu %>% 
#   filter(Admin2 == "Nantucket")

```

```{r}
# jhu %>% 
#   filter(Country_Region == "US") %>% 
#   filter(Province_State == "Virginia") %>% 
#   select(Admin2, Confirmed) %>% 
#   slice_max(Confirmed)
# 
# 
# jhu %>% 
#   filter(Country_Region == "US") %>% 
#   filter(Province_State == "Virginia") %>% 
#   select(Admin2, Deaths) %>% 
#   filter(Deaths > 0) %>% 
#   slice_min(Deaths) 
```


County population

```{r}
# county_pop <- 
#   read_csv("county_pop.csv", skip = 5,
#                        col_names = c("county", "census", "estimates_base", "2010",
#                                      "2011", "2012", "2013", "2014", "2015", "2016",
#                                      "2017", "2018", "2019")) %>% 
#   select(county, pop = `2019`) %>% 
#   mutate(county = str_sub(county, 2)) %>%
#   mutate(county = str_replace(county, " County", "")) 
#   
# county_pop$county[which(str_detect(county_pop$county, "(C|c)ity"))]
  # pluck("county", 90)
  # separate(county, into = c("county", "state"), sep = " County, ") #%>%
  # filter(state == "Virgnia")
  # select(county, state, pop)
county_pop_est <-
  read_csv("co-est2019-alldata.csv") %>% 
    janitor::clean_names() %>% 
    mutate(FIPS = as.numeric(state) * 1000 + as.numeric(county)) %>% 
    # mutate(FIPS = str_c(state, county)) %>% 
    select(FIPS, stname, ctyname, pop = popestimate2019)
```

County land area

```{r}
land_area <- 
readxl::read_xls("LND01.xls") %>% 
  janitor::clean_names() %>%
  rowwise() %>% 
  mutate(land_area = max(c_across(ends_with("d")))) %>% 
  transmute(areaname, FIPS = as.numeric(stcou), 
            land_area)
```

```{r}
land_area %>% 
  summary()

covid_cum %>% 
  filter(is.na(FIPS))
```


```{r}
county_pop_est %>% 
  left_join(land_area) %>% 
  transmute(FIPS, areaname, pop, land_area, density = pop / land_area) %>% 
  summary()
```


```{r}
covid_cum_final <- 
covid_cum %>% 
  left_join(county_pop_est) %>% 
  left_join(land_area) %>% 
  transmute(FIPS, county = Admin2, state = Province_State,
            Confirmed, Deaths, Case.Fatality_Ratio,
            pop, land_area, density = pop / land_area) %>% 
  janitor::clean_names() %>% 
  drop_na() 

covid_cum_final %>% 
  filter(density < 5000) %>% 
  ggplot(aes(x = density, y = case_fatality_ratio)) + 
  geom_jitter() + 
  geom_jitter(data = covid_cum_final %>% 
                filter(state == "New York",
                       density < 5000),
                       color = "red")

```

```{r}
covid_cum_final %>% 
  filter(case_fatality_ratio > 10) %>% 
  arrange(state) %>% 
  count(state)
```


```{r}
covid_cum_final %>% 
  filter(state == "Virginia",
         county %in% c("Amelia", "Buckingham", "Charlotte", 
                       "Cumberland", "Lunenberg", "Nottoway", "Prince Edward")) %>% 
  summarize(rate = sum(confirmed)/ sum(pop) * 100)
```

```{r}
covid_cum_final %>% 
  filter(state == "Virginia") %>% 
  pivot_longer(c("confirmed", "deaths"), names_to = "type", values_to = "count") %>% 
  ggplot(aes(x = density, y = count, color = type)) +
  geom_jitter() + 
  geom_smooth()
```

```{r}
covid_cum_final %>% 
  filter(state == "Virginia") %>% 
  arrange(desc(deaths))

covid_cum_final %>% 
  filter(state == "Virginia") %>% 
  arrange(desc(confirmed))

covid_cum_final %>% 
  filter(state == "Virginia") %>% 
  arrange(desc(case_fatality_ratio))
```


```{r}
options(scipen = 999)
covid_cum_final %>% 
  filter(state == "Virginia") %>% 
  mutate(fairfax = ifelse(
    county %in% c("Fairfax", "Henrico", "Prince William", "Arlington", "Loudon"), 
    1, 0),
         ) %>% 
  lm(deaths ~ confirmed * fairfax + density, .) %>% 
  summary()
```

```{r}
covid_cum_final %>% 
  filter(state == "Virginia",
         confirmed < 2500) %>% 
  ggplot(aes(x = confirmed, y = deaths)) +
  geom_jitter() + 
  geom_smooth(method = "lm")
```

```{r}
covid_cum_final %>% 
  # filter(state %in% c("New York", "Florida")) %>% 
  # mutate(new_york = ifelse(state == "New York", 1, 0)) %>% 
  lm(deaths ~ confirmed + density + I(density^2), .) %>% 
  summary()
```

```{r}
covid_cum_final %>% 
  filter(state %in% 
           c("New York", "Illinois", "New Jersey", "Massachusetts",
             "Pennsylvania", "Connecticut", "Ohio")) %>% 
  # filter(deaths < 15000) %>% 
  ggplot(aes(x = confirmed, y = deaths, color = state)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = F)  
  # coord_cartesian(xlim = c(0, 10000), ylim = c(0, 1000)) +
  # scale_y_log10()
```

```{r}
covid_cum_nest <- 
covid_cum_final %>% 
  group_by(state) %>% 
  nest()
```


```{r}
covid_cum_nest %>% 
  mutate(
    fit = map(~lm(deaths ~ confirmed + density, data = .)))
```

```{r}
covid_cum_nest %>% 
  # nest(-state) %>% 
  mutate(
    fit = map(data, ~lm(deaths ~ confirmed,  data = .x)),
    tidied = map(fit, broom::tidy),
    glanced = map(fit, broom::glance),
    augmented = map(fit, broom::augment)
    ) %>% 
  unnest(tidied) %>% 
  select(state, term, estimate, std.error, statistic) %>% 
  pivot_wider(names_from = term,
              values_from = c(estimate, std.error, statistic)) %>% 
  arrange(desc(estimate_confirmed))
```

```{r}
covid_cum_final %>% 
  nest(-state)
```

```{r}
covid_cum %>% 
  filter(Province_State == "Rhode Island")
```


```{r}
read_csv("covid-law-dates.csv") %>% 
  janitor::clean_names() %>% 
  select(state = province_state_name,
         contains("policy")
         ) %>% 
  filter(policy_name == "StayAtHome", 
         policy_mandatory_code == 0)
```

