---
title: "Covid Project Regressions"
author: "Tony Carilli"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Data

## Stringency Mobility and Social Capital

```{r}
df <-
  here::here("data", "covid_state_level.csv") %>%
  read_csv() %>%
  group_by(state) %>%
  mutate(
    stringency_dummy = ifelse(stringency_index > 50, 1, 0),
    state_level_dummy = ifelse(state_level_index > 0, 1, 0),
    family_unity_dummy = ifelse(family_unity > 0, 1, 0),
    family_interaction_dummy = ifelse(family_interaction > 0, 1, 0),
    social_support_dummy = ifelse(social_support > 0, 1, 0),
    community_health_dummy = ifelse(community_health > 0, 1, 0),
    institutional_health_dummy = ifelse(institutional_health > 0, 1, 0),
    collective_efficacy_dummy = ifelse(collective_efficacy > 0, 1, 0),
    philanthropic_health_dummy = ifelse(philanthropic_health > 0, 1, 0),
    retail_recreation = zoo::rollmean(
      retail_and_recreation_percent_change_from_baseline,
      7,
      NA,
      "right"
    ),
    grocery_pharmacy = zoo::rollmean(
      grocery_and_pharmacy_percent_change_from_baseline,
      7,
      NA,
      "right"
    ),
    parks = zoo::rollmean(parks_percent_change_from_baseline, 7, NA, "right"),
    transit_stations = zoo::rollmean(
      transit_stations_percent_change_from_baseline,
      k = 7,
      fill = NA,
      align = "right"
    ),
    workplaces = zoo::rollmean(workplaces_percent_change_from_baseline, 7, NA, "right"),
    residential = zoo::rollmean(residential_percent_change_from_baseline, 7, NA, "right"),
    deaths_ma = zoo::rollmean(deaths_new, 7, NA, "right"),
    cases_ma = zoo::rollmean(cases_new, 7, NA, "right"),
    state = str_sub(state, -2),
    stringency_lag_7 = dplyr::lag(stringency_index, 7),
    stringency_lag_14 = dplyr::lag(stringency_index, 14),
    
  ) %>%
  ungroup()

```


## Social Capital Project data

```{r}
state_index <- 
  here::here("data", "social-capital-project-social-capital-index-data .xlsx") %>% 
  readxl::read_xlsx(sheet = "State Benchmarks", skip = 2) %>% 
  janitor::clean_names() %>% 
  select(-starts_with("x"))
```

## state_index subset

```{r}
names(state_index)

state_index_subset <-
  state_index %>%
  select(
    state = state_abbreviation,
    percent_in_fair_or_poor_health,
    percent_rural,
    percent_black,
    percent_without_health_insurance,
    median_household_income,
    percent_adults_graduated_high_school,
    percent_adults_with_ba,
  ) %>%
  mutate(
    poor_health_dummy = ifelse(percent_in_fair_or_poor_health < 16.5, 1, 0),
    health_insurance_dummy = ifelse(percent_without_health_insurance < 10.7, 1, 0)
  )
df %>%
  left_join(state_index_subset)

```


## governor affiliation data

```{r}

governors <-
  read_csv("data/state_politics.csv", skip = 2) %>%
  slice(2:52) %>%
  mutate(
    state = state.abb[match(Location, state.name)],
    state = ifelse(Location == "District of Columbia", "DC", state),
    .before = Location
  ) %>%
  janitor::clean_names() %>%
  select(state, governor_political_affiliation)

```




# Transit case elasticity

```{r}
df %>% 
  plm::plm(transit_stations ~ log(dplyr::lag(cases_new, 7) + 1), 
           index =  c("state", "date"), data = ., model = "within") %>% 
  summary()
```

# Transit death elasticity

```{r}
df %>% 
  plm::plm(transit_stations ~ log(dplyr::lag(deaths_new, 7) + 1), 
           index =  c("state", "date"), data = ., model = "within") %>% 
  summary()
```


```{r}
df %>% 
  mutate(
  confirmed_cases_ma_7 = zoo::rollmean(
      confirmed_cases,
      k = 7,
      fill = NA,
      align = "right"),
  stringency_index_scaled = scale(stringency_index)) %>% 
  plm::plm(residential_percent_change_from_baseline ~ log(dplyr::lag(confirmed_cases_ma_7, 7) + 1) +
             stringency_index_scaled, 
           index =  c("state", "date"), data = ., model = "within") %>% 
  summary()
```





```{r}
df %>% 
  group_by(state) %>%
  filter(!is.na(stringency_index)) %>% 
  summarize(date = max(date)) %>% 
  summarize(date = min(date)) %>% 
  pull(date)
```

# MODEL 1 We LIKE


```{r model-1-we-like}
df %>%
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_dummy * state_level_dummy,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```

```{r mobility-and-social-capital}
df %>%
  plm::plm(
    residential ~ stringency_index * state_level_index + 
      dplyr::lag(cases_ma,1),
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```
## IV model


```{r model-we-like-with-iv}
df %>%
  left_join(governors) %>% 
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_index * state_level_index | 
      stringency_index + stringency_index:state_level_index + 
      governor_political_affiliation + stringency_lag_7 * state_level_index +
      dplyr::lag(cases_new, 1),
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```

```{r}
df %>% 
  group_by(state) %>% 
  mutate(cases_ma_lag_7 = dplyr::lag(cases_ma, 7)) %>% 
  plm::plm(stringency_index ~ state_level_index + cases_ma_lag_7,
           data = .,
           index = "date", 
           model = "within") %>% 
  summary()
```

```{r}
df %>% 
  left_join(state_index, by = c("state" = "state_abbreviation")) %>% 
  filter(date == "2020-11-01") %>% 
  lm(state_level_index ~ median_household_income + percent_adults_with_ba +
             percent_black * percent_rural,
           data = .) %>% 
  summary()
           
  
  
```

```{r final-iv-plm-model}
df %>%
  left_join(governors) %>% 
  left_join(state_index, by = c("state" = "state_abbreviation")) %>% 
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_index * state_level_index | 
      stringency_index + stringency_index:state_level_index + 
      governor_political_affiliation + stringency_lag_7 * state_level_index +
      median_household_income + percent_adults_with_ba +
             percent_black * percent_rural + 
      dplyr::lag(cases_new, 1),
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```

```{r}
df %>%
  left_join(governors) %>% 
  left_join(state_index, by = c("state" = "state_abbreviation")) %>% 
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_index * state_level_index + 
             median_household_income + percent_adults_with_ba +
             percent_black * percent_rural + population_density,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```


```{r}
df %>%
  left_join(governors) %>% 
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ governor_political_affiliation * state_level_dummy,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```


```{r}
df %>%
  left_join(governors) %>% 
  group_by(state) %>%
  mutate(governor = ifelse(governor_political_affiliation == "Republican", 1, 0)) %>%
  ungroup() %>% 
  lm(state_level_dummy ~ governor_political_affiliation,
    data = .,
    ) %>%
  summary()
```
## Race vs family unity

```{r}
state_index %>% 
  left_join(df) %>% 
  lm(family_prosperity_index ~ percent_black, .) %>% 
  summary()
```


## Cross Sectional model

```{r}
df %>% 
  left_join(governors) %>% 
  left_join(state_index, by = c("state" = "state_abbreviation")) %>% 
  filter(date == "2020-11-01") %>% 
  lm(cases_new ~ stringency_index * state_level_index, 
     data = .) %>% 
  jtools::summ()
```
```{r}
df %>% 
  left_join(governors) %>% 
  left_join(state_index) %>% 
  filter(date == "2020-11-01") %>% 
  pull(percent_in_fair_or_poor_health)
```



```{r}
df %>%
  plm::plm(
    cases_new ~ stringency_dummy * state_level_dummy,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```



```{r}
df %>%
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_dummy * state_level_index,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```

```{r}
df %>%
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_dummy * state_level_dummy + 
      population_density,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```

Disaggregated Social Indicators

```{r}
df %>%
  plm::plm(
    cases_new ~ stringency_dummy * state_level_dummy + population_density,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 broom::tidy()

df %>%
  plm::plm(
    cases_new ~ stringency_dummy * family_unity_dummy + population_density,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 broom::tidy()

df %>%
  plm::plm(
    cases_new ~ stringency_dummy * community_
    _dummy + population_density,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 broom::tidy()


```

```{r many-models}
models <-
  list(
    state = 'cases_new ~ stringency_dummy * state_level_dummy + population_density',
    family_unity = 'cases_new ~ stringency_dummy * family_unity_dummy + population_density',
    family_interaction = 'cases_new ~ stringency_dummy * family_interaction_dummy + population_density',
    social_support = 'cases_new ~ stringency_dummy * social_support_dummy + population_density',
    community_health = 'cases_new ~ stringency_dummy * community_health_dummy + population_density',
    institutional_health = 'cases_new ~ stringency_dummy * institutional_health_dummy + population_density',
    collective_health = 'cases_new ~ stringency_dummy * collective_efficacy_dummy + population_density',
    philanthropic_health = 'cases_new ~ stringency_dummy * philanthropic_health_dummy + population_density'
  )

model_frame <- tibble(model = models) %>% 
  mutate(model_name = names(model))

model_frame <- model_frame %>% 
  mutate(model = map(model, as.formula))

estimated_models <- 
  model_frame %>% 
  mutate(fit = map(model, ~plm::plm(., data = df, index = c("date"), model = "within")))

models_tidy <- 
  estimated_models %>% 
  transmute(model_name, betas = purrr::map(fit, broom::tidy))

models_tidy %>% 
  pivot_wider(names_from = model_name,
              values_from = betas) %>% 
  pluck("state")
  

estimated_models$fit %>% 
  purrr::map_df(broom::tidy) %>% 
  mutate(model = rep(names(models), each = 4), .before = "term") %>% 
  pivot_longer(cols = 3:6,
               names_to = "estimate",
               values_to = "value")
  

estimated_models$fit %>% 
  purrr::map_df(broom::tidy) %>% 
  mutate(model = rep(names(models), each = 4), .before = "term") %>% 
  pivot_wider(names_from = model)

```

```{r}
models_tidy %>% 
  unnest_wider(betas) %>% 
  # unnest_longer(term) %>% 
  unnest_wider(estimate, names_sep = "_")
```


# fair 
interacted with social capital

```{r}
df %>% 
  left_join(state_index_subset) %>% 
  plm::plm(
    cases_new ~ poor_health_dummy * state_level_dummy ,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
  

df %>% 
  left_join(state_index_subset) %>% 
  plm::plm(
    cases_new ~ health_insurance_dummy * state_level_dummy * stringency_dummy,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
```
# increase stringency dummy to .65

```{r}
df %>% 
  mutate(stringency_dummy = ifelse(stringency_index > 65, 1, 0)) %>% 
  left_join(state_index_subset) %>% 
  plm::plm(
    cases_new ~ stringency_dummy * state_level_dummy,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
```
# stringency lagged

```{r}
df %>% 
  left_join(state_index_subset) %>% 
  plm::plm(
    cases_new ~ stringency_lag_7 * state_level_index,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()

df %>% 
  left_join(state_index_subset) %>% 
  plm::plm(
    cases_new ~ stringency_lag_14 * state_level_index,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
```

# estimate stringency

```{r}
df %>% 
  left_join(state_index_subset) %>% 
  plm::plm(
    cases_new ~ stringency_lag_14 * state_level_index,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
```


# add governors political affiliation

## looking for instruments

```{r}
df %>% 
  left_join(state_index_subset) %>% 
  left_join(governors) %>% 
  mutate(deaths_lag = dplyr::lag(deaths_ma, 7),
         cases_lag = dplyr::lag(cases_ma, 14),
         stringency_lag = dplyr::lag(stringency_index, 14)) %>% 
  plm::plm(
    stringency_index ~  percent_without_health_insurance + 
      governor_political_affiliation +  percent_rural + 
      + median_household_income + population_density +
      percent_black + percent_in_fair_or_poor_health +
      percent_adults_with_ba +
      deaths_lag + cases_lag + stringency_lag,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
```
## governor affiliation - percent rural

```{r}
df %>% 
  left_join(state_index_subset) %>% 
  left_join(governors) %>% 
  # filter(date == "2020-11-01") %>% 
  # select(percent_rural, governor_political_affiliation) %>% 
  mutate(republican = ifelse(governor_political_affiliation == "Republican", 1, 0)) %>% 
  lm(
    republican ~ percent_rural,
    data = .
  ) %>%
 summary()
```

```{r}
df %>% 
  left_join(state_index_subset) %>%
  left_join(governors) %>% 
  plm::plm(stringency_index ~ percent_rural + governor_political_affiliation,
           data = .,
           index = "date",
           model = "within") %>% 
  summary()
```


```{r}
df %>% 
  left_join(state_index_subset) %>% 
  left_join(governors) %>% 
  plm::plm(
    cases_new ~ stringency_lag_14 * state_level_index + 
      percent_rural,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
 summary()
```

---

```{r}
df %>%
  group_by(state) %>%
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 14)) %>%
  ungroup() %>% 
  plm::plm(
    cases_new ~ stringency_dummy * state_level_index + 
      population_density,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
```

```{r}
df %>% 
  mutate(pop_over_50 = (population_age_50_59 + population_age_60_69 + 
                          population_age_70_79 + population_age_80_and_older) / population) %>% 
  plm::plm(
    cases_new ~ stringency_dummy * state_level_index + 
      population_density + pop_over_50,
    data = .,
    index = c("date"),
    model = "within"
  ) %>%
  summary()
  
```



```{r}
df %>% 
  lm(log(cases_new + 1) ~ state_level_index * stringency_dum + date +
       population_density, .) %>% 
  jtools::summ()
  
```


```{r}
df %>% 
  group_by(state) %>% 
  mutate(lag_stringency_index = dplyr::lag(stringency_index, 7)) %>% 
  lm(deaths_new ~ state_level_index +  population_density + lag_stringency_index, .) %>% 
  jtools::summ(digits = 4, scale = TRUE)
```

```{r}
df %>% 
  group_by(state) %>% 
  filter(!is.na(transit_stations_percent_change_from_baseline)) %>% 
  mutate(elasticity = log(transit_stations_percent_change_from_baseline + 100) - 
           dplyr::lag(log(transit_stations_percent_change_from_baseline + 100), 1) /
           (log(cases_new + 1) - log(dplyr::lag(cases_new, 1)) + 1)
  ) %>% 
  summary()
```

```{r}
df %>%
  group_by(state) %>%
  filter(!is.na(transit_stations_percent_change_from_baseline),
         confirmed_cases > 0) %>%
  mutate(
    confirmed_cases_ma_7 = zoo::rollmean(
      confirmed_cases,
      k = 7,
      fill = NA,
      align = "right"
    ),
    transit_stations_percent_change_from_baseline_ma = zoo::rollmean(
      transit_stations_percent_change_from_baseline,
      7,
      align = "right",
      fill = NA
    ),
    percent_change_cases = log(confirmed_cases_ma_7) - dplyr::lag(log(confirmed_cases_ma_7), 7),
    percent_change_mobility = log(transit_stations_percent_change_from_baseline_ma + 100) -
      dplyr::lag(
        log(transit_stations_percent_change_from_baseline_ma + 100),
        1
      ),
    elasticity = percent_change_mobility / percent_change_cases
  ) %>%
  filter(is.finite(elasticity)) %>% 
  ggplot(aes(x = state, y = elasticity)) +
  geom_boxplot() # %>%
  # select(elasticity) %>% 
  # summarize(mean(elasticity, na.rm = TRUE))
  # filter(state == "US_AZ") %>%
  # select(date,
  #        confirmed_cases,
  #        transit_stations_percent_change_from_baseline,
  #        confirmed_cases_ma_7,
  #        transit_stations_percent_change_from_baseline_ma,
  #        elasticity
  #        ) %>%
  # filter(elasticity > 4)
```

```{r}
df %>% 
  group_by(state) %>% 
  filter(confirmed_cases > 0) %>% 
  mutate(mov_avg = zoo::rollmean(confirmed_cases, k = 7, fill = NA, align = "right"))

zoo::rollmean(x = df$confirmed_cases, k = 7, fill = NA, align = "right")
```

```{r}
df %>%
  filter(state == "US_AZ") %>%
  mutate(
    transit_stations_percent_change_from_baseline_ma = zoo::rollmean(transit_stations_percent_change_from_baseline,
                                                                     7, align = "right", fill = NA)) %>%
  ggplot(aes(x = date, y = transit_stations_percent_change_from_baseline_ma)) +
  geom_line()
```

```{r warning=FALSE}

df %>% 
  mutate(state = str_sub(state, -2)) %>% 
  filter(state != "VI") %>% 
  group_by(state) %>% 
  filter(!is.na(deaths_new)) %>% 
  mutate(cases_new_ma = zoo::rollmean(cases_new, k = 7, fill = NA, align = "right")) %>% 
  ggplot(aes(x = date, y = cases_new_ma)) + 
  geom_line() +
  geofacet::facet_geo(~ state, 
                      grid = "us_state_grid2"
    
  )
  
```

```{r}
df %>% 
  filter(date == "2020-11-01") %>% 
  lm(confirmed_cases ~ state_level_index + family_unity + social_support +
       ehhi + rhhi +  
       population_density + stringency_index, .) %>% 
  jtools::summ(scale = TRUE, transfrom.response = TRUE, robust = "HC0") 
```

```{r}
df %>% 
  # filter(date == "2020-11-01") %>% 
  lm(confirmed_cases ~ state_level_index + family_unity + social_support +
       ehhi + rhhi +
       population_density + stringency_index + 
       state + date, .) %>% 
  jtools::summ(scale = TRUE, transfrom.response = TRUE, robust = "HC0") 
```

```{r}
df %>% 
  filter(date == "2020-11-01") %>% 
  lm(residential_percent_change_from_baseline ~ state_level_index + family_unity + social_support +
       ehhi + rhhi +  
       population_density + stringency_index, .) %>% 
  jtools::summ(scale = TRUE, transfrom.response = TRUE, robust = "HC0") 
```

