---
title: "Mobility EDA"
author: "Carilli"
date: "`r lubridate::today()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: zenburn
    code_folding: hide
    df_print: paged
    fig_caption: true
params:
  fips: 51041
  county: Chesterfield
  state: Virginia
  city: Richmond
  apple_county: Chesterfield County
  state_abb: VA
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, options(scipen = 999, digits = 3),
                      message = FALSE, warning = FALSE)
library(tidyverse)
library(ggrepel)
library(directlabels)
library(gtrendsR)
library(here)
```

# Device Exposure Data {.tabset .tabset-fade .tabset-pills}

## County Data

```{r dex-data}
county_dex <- 
  here("data", "county_dex.csv") %>% 
  read_csv() %>% 
  # coerce county to numeric fips code
  mutate(fips = as.numeric(county), .before = county,
  # coerce date to date from character vector
         date = as.Date(date)) %>% 
  select(-county)

# county_dex
```

## State data

```{r state-dex-data}
state_dex <- 
here("data", "state_dex.csv") %>% 
  read_csv() 

state_dex_long <- 
state_dex %>% 
  select(-c(contains("_a"), contains("num"))) %>% 
  # glimpse()
  pivot_longer(dex_income_1:dex_race_white,
               names_to = "category",
               values_to = "cat_dex") 
  
```


## Dex data plots

### `r params$state` 

```{r state-dex-plot}
state_dex %>%
  filter(state == params$state_abb) %>%
  mutate(mov_avg_7 = zoo::rollmean(dex, 7, NA)) %>%
  ggplot(aes(x = date, y = mov_avg_7)) +
  geom_line()
```


### `r params$state` by race

```{r dex-state-plot-race}
state_dex_long %>% 
  filter(state == params$state_abb) %>%
  filter(str_detect(category, "race")) %>% 
  group_by(category) %>% 
  mutate(mov_avg_7 = zoo::rollmean(cat_dex, 7, NA)) %>% 
  ggplot(aes(x = date, y = mov_avg_7, color = category)) + 
  geom_line() 

```

### `r params$state` by income

```{r dex-state-plot-income}
state_dex_long %>% 
  filter(state == params$state_abb) %>%
  filter(str_detect(category, "income")) %>% 
  group_by(category) %>% 
  mutate(mov_avg_7 = zoo::rollmean(cat_dex, 7, NA)) %>% 
  ggplot(aes(x = date, y = mov_avg_7, color = category)) + 
  geom_line() 
```

### `r params$state` by education

```{r dex-state-plot-education}
state_dex_long %>% 
  filter(state == params$state_abb) %>%
  filter(str_detect(category, "education")) %>% 
  group_by(category) %>% 
  mutate(mov_avg_7 = zoo::rollmean(cat_dex, 7, NA)) %>% 
  ggplot(aes(x = date, y = mov_avg_7, color = category)) + 
  geom_line() 

```


### `r params$county` County

```{r dex-county-plot, echo=FALSE}
county_dex %>% 
  filter(fips == params$fips) %>% 
  mutate(mov_avg_7 = zoo::rollmean(dex, 7, NA)) %>% 
  ggplot(aes(x = date, y = mov_avg_7)) +
  geom_line(color = "blue") + 
  # geom_vline(xintercept = as.Date("2020-03-30"), color = "red", size = 1.25) +
  labs(x = NULL,
       y = "7 Day Moving Average",
       title = "Device Exposure Index",
       subtitle = glue::glue({params$county}, " County, ", {params$state}))
```

# Google mobility data {.tabset .tabset-fade .tabset-pills}

## Us by county data wide

```{r google-data-us}
mobility_us_by_county <-
  here("data", "us_mobility_report.csv") %>%
  read_csv(
    col_types =
      cols(
        country_region_code = col_character(),
        country_region = col_character(),
        sub_region_1 = col_character(),
        sub_region_2 = col_character(),
        metro_area = col_character(),
        iso_3166_2_code = col_character(),
        census_fips_code = col_integer(),
        date = col_date(format = ""),
        retail_and_recreation_percent_change_from_baseline = col_double(),
        grocery_and_pharmacy_percent_change_from_baseline = col_double(),
        parks_percent_change_from_baseline = col_double(),
        transit_stations_percent_change_from_baseline = col_double(),
        workplaces_percent_change_from_baseline = col_double(),
        residential_percent_change_from_baseline = col_double()
      )
  ) %>% 
  # filter(country_region_code == "US") %>%
  select(date,
         state = sub_region_1, county = sub_region_2, fips = census_fips_code,
         retail_and_recreation_percent_change_from_baseline:
           residential_percent_change_from_baseline) 
  

# mobility_us_by_county
```

## US by county tidied

```{r google-data-long-county, warning=FALSE}
mobility_us_by_county_long <- 
mobility_us_by_county %>%
  filter(!is.na(fips)) %>%
  pivot_longer(
    -c(date,
       state,
       county,
       fips),
    names_to = "measure",
    values_to = "percent_change_from_baseline",
    names_pattern = "(.*)_percent_change_from_baseline"
  ) 
# mobility_us_by_county_long
```
Some counties proved problematic due to missing observations.  We'll have to creating moving averages as we need paricular counties.

## US by State tidied

```{r google-data-long-state, message=FALSE, warning=FALSE}
mobility_us_by_state_long <- 
mobility_us_by_county %>% 
  filter(!is.na(state), is.na(fips)) %>% 
  select(-c(county, fips)) %>% 
  pivot_longer(-c(date, state), names_to = "measure", 
               values_to = "percent_change_from_baseline",
               names_pattern = "(.*)_percent_change_from_baseline") %>% 
  group_by(measure, state) %>% 
  mutate(mov_avg_7 = zoo::rollmean(percent_change_from_baseline, 7, fill = "NA"))

# mobility_us_by_state_long


```

## US tidied

```{r google-data-long-us, warning=FALSE}
mobility_us_long <- 
mobility_us_by_county %>% 
  filter(is.na(state)) %>% 
  select(-c(state, county, fips)) %>% 
  pivot_longer(-date, names_to = "measure", 
               values_to = "percent_change_from_baseline",
               names_pattern = "(.*)_percent_change_from_baseline") %>% 
  group_by(measure) %>% 
  mutate(mov_avg_7 = zoo::rollmean(percent_change_from_baseline, 7, fill = "NA")) 

# mobility_us_long
```

## Google Mobility Plots

### US plot

```{r us-mobility-plot}
mobility_us_long %>% 
  group_by(date) %>% 
  ggplot(aes(x = date, y = mov_avg_7, color = measure)) + 
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = measure), method = "smart.grid") +
  labs(x = NULL,
       y = "7 Day Moving Average",
       caption = "Source: COVID-19 Community Mobility Reports",
       title = "US Mobility",
       subtitle = "Measured as percent change from baseline")
  
```

### State plot `r params$state`

```{r state-mobility-plot}
mobility_us_by_state_long %>% 
  filter(state == params$state) %>% 
  ggplot(aes(x = date, y = mov_avg_7, color = measure)) + 
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = measure), method = "smart.grid") +
  labs(x = NULL,  
       y = "7 Day Moving Average",
       caption = "Source: COVID-19 Community Mobility Reports",
       title = glue::glue({params$state}, "Mobility", .sep = " "),
       subtitle = "Measured as percent change from baseline")
         
```

### `r params$county` plot

```{r google-county-plot}
mobility_us_by_county_long %>% 
  filter(fips == params$fips) %>% 
  group_by(measure) %>% 
  mutate(mov_avg_7 = zoo::rollmean(percent_change_from_baseline, 7, fill = "NA")) %>% 
  ggplot(aes(x = date, y = mov_avg_7, color = measure)) + 
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = measure), method = "smart.grid") +
  labs(x = NULL,  
       y = "7 Day Moving Average",
       caption = "Source: COVID-19 Community Mobility Reports",
       title = glue::glue({params$county}, {params$state}, "Mobility", .sep = " "),
       subtitle = "Measured as percent change from baseline")

  
```

# Apple mobility data {.tabset .tabset-fade .tabset-pills}

We can apple mobility data by the US, by state, by county, and by city.

## Data

### US

Driving, walking, and transit for US total

```{r apple-us}
apple_us <- 
here("data", "applemobilitytrends-2020-08-05.csv") %>%
  read_csv(
    col_types = cols(
      .default = col_double(),
      geo_type = col_character(),
      region = col_character(),
      transportation_type = col_character(),
      alternative_name = col_character(),
      `sub-region` = col_character(),
      country = col_character(),
      `2020-05-11` = col_double(),
      `2020-05-12` = col_double()
    )
  ) %>%
  filter(region == "United States") %>% 
  select(-c(geo_type, region, alternative_name, `sub-region`, country)) %>% 
  pivot_longer(-transportation_type,
               names_to = "date",
               values_to = "index")
# apple_us  

```

### State level

The only data available by state are driving data.

```{r apple-state}
apple_state <- 
here("data", "applemobilitytrends-2020-08-05.csv") %>%
  read_csv(
    col_types = cols(
      .default = col_double(),
      geo_type = col_character(),
      region = col_character(),
      transportation_type = col_character(),
      alternative_name = col_character(),
      `sub-region` = col_character(),
      country = col_character(),
      `2020-05-11` = col_double(),
      `2020-05-12` = col_double()
    )
  ) %>%
  select(-c(alternative_name)) %>% 
  filter(region %in% state.name) %>% 
  pivot_longer(starts_with("2020"),
               names_to = "date",
               values_to = "index") %>% 
  select(date, state = region, transportation_type, index)

# apple_state
```

### County level

Driving is the only type available at the county level.

```{r apple-county}
apple_county <- 
here("data", "applemobilitytrends-2020-08-05.csv") %>%
  read_csv(
    col_types = cols(
      .default = col_double(),
      geo_type = col_character(),
      region = col_character(),
      transportation_type = col_character(),
      alternative_name = col_character(),
      `sub-region` = col_character(),
      country = col_character(),
      `2020-05-11` = col_double(),
      `2020-05-12` = col_double()
    )
  ) %>%
  filter(geo_type == "county", country == "United States") %>% 
  select(-c(alternative_name, geo_type, country), county = region, state = `sub-region`,
         starts_with("2020")) %>% 
  pivot_longer(starts_with("2020"),
               names_to = "date",
               values_to = "index")
# apple_county
```

### City Level

```{r apple-city}
apple_city <- 
here("data", "applemobilitytrends-2020-08-05.csv") %>%
  read_csv(
    col_types = cols(
      .default = col_double(),
      geo_type = col_character(),
      region = col_character(),
      transportation_type = col_character(),
      alternative_name = col_character(),
      `sub-region` = col_character(),
      country = col_character(),
      `2020-05-11` = col_double(),
      `2020-05-12` = col_double()
    )
  ) %>%
  filter(geo_type == "city", country == "United States") %>% 
  select(-c(alternative_name, geo_type, country), city = region, state = `sub-region`,
         starts_with("2020")) %>%
  pivot_longer(starts_with("2020"),
               names_to = "date",
               values_to = "index")

# apple_city
```

## Apple Plots

### US

```{r apple-us-plot}
apple_us %>% 
  group_by(transportation_type) %>% 
  mutate(mov_avg_7 = zoo::rollmean(index, 7, NA)) %>% 
  ggplot(aes(x = as.Date(date), y = mov_avg_7, color = transportation_type)) + 
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = transportation_type), method = "smart.grid") +
  labs(x = NULL,  
       y = "7 Day Moving Average",
       caption = "Source: Apple Mobility Trends",
       title = "US Mobility",
       subtitle = "Index (100 basline)")
  
```

### `r params$state` Mobility Plot

```{r apple-state-plot}
apple_state %>% 
  filter(state == params$state) %>% 
  mutate(mov_avg_7 = zoo::rollmean(index, 7, NA)) %>%
  ggplot(aes(x = as.Date(date), y = mov_avg_7, color = transportation_type)) +
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = transportation_type), method = "smart.grid") +
  labs(x = NULL, 
       y = "7 Day Moving Average",
       caption = "Source: Apple Mobility Trends",
       title = glue::glue({params$state}, "Mobility", .sep = " "),
       subtitle = "Index (100 basline)")
```

### `r params$county`, `r params$state` Mobility

```{r apple-county-plot}
apple_county %>% 
  filter(state == params$state, county == params$apple_county) %>% 
  mutate(mov_avg_7 = zoo::rollmean(index, 7, NA)) %>%
  ggplot(aes(x = as.Date(date), y = mov_avg_7, color = transportation_type)) +
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = transportation_type), method = "smart.grid") +
  labs(x = NULL, 
       y = "7 Day Moving Average",
       caption = "Source: Apple Mobility Trends",
       title = glue::glue({params$apple_county}, {params$state}, "Mobility",
                          .sep = " "),
       subtitle = "Index (100 basline)")
```

### `r params$city`, `r params$state` Mobility

```{r}
apple_city %>% 
  filter(city == "Richmond") %>% 
  mutate(mov_avg_7 = zoo::rollmean(index, 7, NA)) %>%
  ggplot(aes(x = as.Date(date), y = mov_avg_7, color = transportation_type)) +
  geom_line(show.legend = FALSE) +
  directlabels::geom_dl(aes(label = transportation_type), method = "smart.grid") +
  labs(x = NULL, 
       y = "7 Day Moving Average",
       caption = "Source: Apple Mobility Trends",
       title = glue::glue({params$city}, ", ", {params$state}, " Mobility"),
       subtitle = "Index (100 basline)")
  
```


# Appendix FIPS

```{r appendix, eval=FALSE}
library(rvest)
library(listviewer)
web_page <- 
  read_html("https://www.nrcs.usda.gov/wps/portal/nrcs/detail/?cid=nrcs143_013697",
            as.data.frame = TRUE, stringsAsFactors = TRUE) 

tbls <- 
html_nodes(web_page, "table")

head(tbls)

tbls[[6]][1]

tbls_ls <- 
web_page %>%
  html_nodes("table")

fips <- 
tbls_ls %>% as.data.frame()

fips %>% 
  slice(31:3261) %>% 
  select(fips = X1, county = X2, state = X3) %>% 
  write_csv(here("data", "fips.csv"))
```

```{r fips-lookup}
here("data", "fips.csv") %>% 
  read_csv() %>% 
  DT::datatable()
```

# Appendix Cities

```{r cities-lookup}
apple_city %>% 
  distinct(city, state) %>% 
  DT::datatable()
```

