---
title: "Social Capital and Covid-19"
author: "Carson"
date: '`r format(lubridate::today(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    theme: cerulean
    highlight: tango
    code_folding: hide
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Social Capital and Covid-19

## I Introduction
  
In light of the on-going Covid-19 pandemic, scholars are beginning to examine the relationship between social capital and the spread and prevention of Covid-19 (Wong and Kohler, 2020; Borgonovi and Andrieu, 2020; Borgonovi, Andrieu, and Subramanian, 2020; Wu, 2020; Wu et al., 2020; Bartscher et al., 2020; Durante et al., 2020).^[Putnam (2000: 19) defines social capital as the "connections among individuals---social networks and the norms of reciprocity and trustworthiness that arise from them."] The influence social capital has on health outcomes in general is well-studied (Putnam, xxxx), and there is even some recognition that social capital influences the spread of infectious diseases (H1N1, xxxx). While the emerging literature on social capital and Covid-19 generally shows that social capital encourages preventative behavior, few recognize the extent to which social capital complements formal, governmental public health responses like lockdown and stay-at-home orders.^[Income also influences the effectiveness, i.e., compliance with, lockdown orders (Wright et al., 2020).] Without this recognition, the efficacy of social capital and formal public health rules tend to be overestimated, which can lead to poor public health policy.  

It is plausible that people with stronger social capital networks might have more interactions with each other, which could encourage infectious behavior. It is also plausible that greater social capital encourages preventative behavior via private and communal efforts to prevent infection, adopting new informal rules, and enforcing formal public health rules. Existing evidence suggests the latter effect is larger than the former. People living in counties (in the United States) with more social capital reduced their mobility “earlier and to a higher degree” than people in counties with less social capital (Borgonovi and Andrieu (2020: xx; also, see Borgonovi, Andrieu, and Subramanian, 2020).^[Borgonovi, Andrieu, and Subramanian (2020) show that infection rates were initially higher in counties with more social capital, but transmission and mortality rates declined as people learned more about Covid-19. Similarly, Bartscher et al. (2020) suggests that the number of Covid-19 cases across Europe was initially higher in countries with higher social capital, but that cases increased at a slower rate in countries with higher social capital.] There is also suggestive evidence that people in areas with more social capital are more likely to follow formal public health rules, e.g., wearing face masks (Wu, 2020) and are more likely to have higher rates of testing (Wu et al., 2020).

Our study is the first to argue social capital is complementary to formal public health rules and that it exerts an independent effect on Covid-19 transmission. With state-level data on Covid-19 cases, social capital, and stringency---a measure of the severity of public health measures---we develop a panel model to assess these relative effects. Overall, we find social capital and stringency exert significant, independent effects on daily Covid-19 cases, and that higher social capital makes the effect of stringency larger for a given level of stringency. That is, states that have higher measures of social capital *and* stringency have fewer cases of Covid-19 than states that merely adopt tougher stringency regulations. Furthermore, we find that this result is robust to the inclusion of various control variables, instrumental variables, and alternative measures of social capital.

The primary implication is that social capital plays an identifiable role in how people spread infectious diseases, which should be recognized in subsequent research in economics and public health. Furthermore, whereas Bartscher et al. (2020) suggests social capital becomes less important once lockdown orders are issued, our results suggest that social capital plays a consistently complementary role with more formal rules regarding Covid-19, from from March 1, 2020 to November 1, 2020---and when each month is isolated. 

In addition to the emerging economic literature on Covid-19 and the relationship between social capital and Covid-19, this study also has implications for the larger economic literature on institutions and the extent to which formal and informal rules influence behavior. While many economists understand the role institutions have in creating incentives to engage in commercial activity (North, 1990; Baumol, 1990; Ostrom, 2005; Boettke and Coyne, 2003; Heller, 2009; Williamson, 2009; Williamson, 2011; Williamson and Kerekes, 2011; Storr, 2013; Vachris and Isaacs, 2017), the effects those institutions have on infectious and preventative behaviors---and prevalence rates---are less clear.^[While the *disease versus institutions* debate analyzes some of these issues, the focus is on understanding the determinants of economic development, not the determinants of infectious diseases (Sachs and Maloney, xxxx; Acemoglu and Robinson, xxx; Sachs, xxxx; Acemoglu and Robinson, xxxx; Bhattacharya, xxxx).] 

There is some recognition that when incentives encourage individuals to provide public health related goods and services, they tend to provide those services. For example, coal mining companies provided sanitation services on par with municipalities (Fishback, 1993); groups of workers provided industrial-sickness funds that were effective means of insurance (Murray, 2007); private water companies lowered mortality rates in Argentina (Galiani et al., 2005); textile mills and railroad companies provided mosquito control services prior to significant public health measures (Carson, 2016). Individuals can also alter the incentives and rules they face to provide mosquito-control (Carson, 2020a), and to encourage Covid-19 prevention (Carson, forthcoming). Relatedly, people are better able to avoid infectious diseases and/or lessen mortality rates when they live in areas with greater economic freedom (Geloso and Pavlik, 2020; Geloso and Candela, 2020). The importance of social capital---and how social capital influences disease prevention---provides a complementary approach to these studies, especially as economic freedom is a driver of social capital (Berggren and Jordahl, 2006).

Informal rules or social norms also influence behavior, e.g., pro-environmental behaviors (Farrow et al., 2017) (see, Brennan et al., 2014; and Bichierri, xxxx, for general treatments on social norms). Yet, few recognize the extent to which social norms influence behaviors related to the spread of infectious diseases (see, Carson, 2016, for one attempt regarding HIV in the United States). The present study provides additional insight into how social capital influences disease-related behaviors *becuase of* social norms. That is, with an analysis on the subcomponents of social capital we can better understand which kinds of social norms  might be more or less relevant for disease-related behaviors.

The importance of social capital for disease prevention---relative to other kinds of institutions and on its own---remains debatable. In the 18th and 19th centuries, for instance, formal economic and political institutions like the American Constitution and the federal structure of government made it more difficult to develop public health bureaus---to provide filtered water and respond to disease outbreaks--- especially at the state and federal levels of government. Such institutional constraints protected private property, they fostered entrepreneurial activity, and led to economic development, but they delayed efforts to eliminate cholera and yellow fever (Troesken, 2015) and various diseases that spread primarily within animal and livestock populations (Olmstead and Rhode, 2015). Thus, social capital might be less effective relative to formal bureaucracies of public health that can wield policing power.^[Similar to Bhattacharya's (2008) point that both institutions and diseases influence economic development---but at different times---there might be different times in history where social capital is more important relative to formal institutions of public health and *vice versa*. We do not address such issues in this paper, but any such discussion would complement our analysis nicely.] While there were relatively strong norms regarding individualism in the United States---which suggests greater social capital (Belmann et al., 2018) and a lower willingness to regulate economic activity (Williamson, 2009; Williamson, 2013; Davis and Williamson, 2016; Williamson, 2020; Cline and Williamson, 2019)---individuals did eventually create more permanent and effective public health bureaus. 

Furthermore, Troesken (2015: 15-18) describes and critiques *the ideology of the township*, which suggests that tight-knit communities, e.g., American towns during the late 18th and early 19th centuries, were places where people could  easily communicate with their neighbors and---if a public health issue arose---they could appeal to their neighbor's sense of community to resolve the iss and encourage proper health practices.^[Troesken focuses on Smallpox, which suggests the appropriate public health practice was inoculation or variolation.] However, Troesken argues this kind of social capital is not robust.^[Tullock (1969) refers to a similar idea in the form of *social pressure* as a means to resolve collective action problems.] Our results suggest otherwise. 

A related theme in the economics of institutions literature is that formal rules that encourage economic development are dependent on supporting informal rules. The implication is that mimicking formal rules without the underlying cultural values and informal rules (or *metis* following Boettke, Coyne, and Leeson, 2008) is not appropriate and likely to evince worse outcomes. Evidence on the efficacy of matching formal and informal institutions suggests that areas with higher levels of trust, self-reliance, and civic belief foster economic and political institutions that protect property rights, promote openness, and economic development (Vachris and Isaacs, 2017). On the contrary, government-led attempts to foster social capital or create homogeneity---for example, through language laws or the creation of a civil court---create unintended, harmful effects (Carilli, Coyne, and Leeson, 2008). 

We are hesitant about attempts to alter the level and/or kind of social capital so as to influence covid-19 transmission rates. Social capital is not easily alterable as if it were a lever one person or a large group of people can toggle. Even though social capital bolsters formal rules, this does not necessarily suggest policymakers should target areas with less social capital, let alone double-down along the intensive margin (as suggested by Bartscher et al., 2020). Even though areas with less social capital will not experience the complementary effect social capital has on stringency, those areas might have fewer interactions in the first place which might leave them with fewer Covid-19 cases. In any event, the relationships between social capital, formal public health rules, and the spread of infectious diseases enhances our understanding of how people spread infectious diseases.

## II Social Capital, Formal Rules, and Covid-19

Social capital takes various forms between individuals so as to encourage cooperation, which then influences health outcomes. Two of the main forms of social capital include *bridging* between people within a social network and *bonding* between people from different social networks. We do not take a stance as to which form of social capital is more or less relevant in the context Covid-19. We do, however, provide evidence below regarding the measured components of social capital. Furthermore, while social capital can be used for legitimate and nefarious activities, we are solely interested in the ways in which social capital influences disease related behaviors. 

Social capital can influence health outcomes in multiple ways (Putnam, 2000): people with greater access to social capital or in a network with stronger social capital might have more tangible assistance from others, e.g., money or health care; they might have more access to organizations that lower the cost of health care; they might have more opportunities to innovate and provide goods and services related to health; they might be in a social network where social norms enforce healthy behaviors, e.g., don't take the last bite of food; they might also have more of an ability to advance a political agenda to alter health policy; and there might a physiological response because of social capital that improves immune systems(Putnam, 2000).

Our argument about the effect social capital has on Covid-19 prevalence rates is conditional; that is, there are conditions under which social capital encourages preventative behavior. If these conditions are met, we expect to see a lower rate of Covid-19 transmission. 

First, the effectiveness of social capital---as a means of limiting the spread of Covid-19 depends on the normative beliefs and attitudes people have about Covid-19 and the value of disease prevention. This follows the general understanding that social capital might encourage legitimate and nefarious activities, depending on people's goals. If people know how Covid-19 spreads and they value its prevention, whatever social capital they might have can be an effective means of prevention. However, some groups might legitimately value other goals more than prevention, e.g., maintaining cultural practices, following religious observances, pursuing economic activity, etc. Thus, merely having social capital does not necessarily encourage preventative behavior. 

Second, the effectiveness of social capital also depends on the extent, quality, kind, or strength of social capital people have. The kinds (and amount) of bonds people form vary in different school groups, business environments, and familial and social interactions; this is why some refer to social capital as having an entrepreneurial quality (Chamlee-Wright, xxxx; Carilli, Coyne, and Leeson (2008). If we can separate groups into those with *strong* and *weak* social capital, for example, we expect the group with stronger social capital to have a larger effect on preventative behavior, given their normative beliefs. 

Table 1 depicts a way to visualize these two conditions. The rows separate groups by normative beliefs and attitudes; generally there are groups that value prevention and those that do not. The columns separate groups into strong and weak kinds of social capital.

```{r}
  tribble(
    ~na, ~Strong_Social_Capital, ~Weal_Social_Capital,
    "Prevention is valuable", "I-Social capital lowers prevalence rates by *a*", "II-Social capital lowers prevalence rates by *b*",
    "Prevention is not", "III-Social capital raises prevalence rates by \"U+03B1", "IV-Social capital raises prevalence rates by b"
  )
```

Thus, the effect social capital might have on Covid-19 prevalence rates varies by quadrant. Looking at the left-side column---where social capital is strong---the magnitude of the effect social capital has on prevalence rates should be larger in the left-side quadrants. People might value preventative behavior or not, but the strong social capital will play a larger role to pursue a given goal than when their social capital is weaker in the right-side quadrants. Specifically, where people 1) value preventative behavior and 2) they have strong social capital, we expect social capital to have the largest, negative effect on Covid-19 prevalence rates (the upper-left quadrant). the effectiveness of social capital likely falls as the strength of social capital falls (the upper-right quadrant). As people want to pursue other goals, aside from preventative behavior, the effect of social capital might encourage them to behave in infectious ways, which will increase Covid-19 prevalence rates. Social capital should increase prevalence rates here (in the lower-left quadrant); weaker social capital will increase prevalence rates at a lesser rate (in the lower-right quadrant.

This framework is somewhat easy to operationalize, especially given somewhat objective measures of social capital, e.g., the Joint Economic Committee's social capital project. This index allows for a distinction between state with *high* social capital and states with *low* social capital. While it is quite difficult to measure where (large) groups of people value anything, let alone Covid-19 prevention, we attempt to capture this idea by accounting for a state's political affiliation, among other control variables.

Two testable implications follow:

* People that value Covid-19 prevention will have lower prevalence rates more when they have stronger social capital. 

* People that do not value Covid-19 prevention will have higher prevalence rates more when they have stronger social capital. 

Given normative beliefs and attitudes that value prevention, we posit social capital has a negative effect on prevalence rates because social capital raises the private costs to individuals of breaking the rules that a group perceives as legitimate, namely formal and informal public health rules. Municipal face mask ordinances and state-wide lockdown orders are examples of the former kind of public health rules, while wearing a face mask whilst in any public setting and covering ones' mouth for a sneeze are examples of the latter kind.^[Whereas formal rules specify a primary directive and secondary qualifiers, a centralized enforcement mechanism, and it is a rule derived from some kind of governing body, informal rules might only specify a primary directive with few secondary qualifiers, enforcement is decentralized, and it is a rule derived from the normative attitudes of the individuals to which the rule applies (Brennan et al., 2014).] While informal rules are difficult to systematically measure in large groups, we focus on the relationship between formal rules and social capital below.^[See Carson (forthcoming) on the relevance of rules to encourage preventative behavior in the context of Covid-19.] 

Social capital encourages rule-following behavior---again assuming people have beliefs and normative attitudes that value preventative behavior---because it raises the private cost to individuals of breaking rules. Given many people believe public health rules are legitimate and valuable, breaking those rules is taken as an affront punishable informally by a personal rebuke, a worsening reputation, a refusal of service, or fewer social contacts. More formal measures of enforcement and punishment are also relevant, e.g., calling a public health department or the police. Relatedly, people are better able to monitor the behavior of others in areas with stronger social capital given the numerous social ties; this suggests that infractions are more likely to be discovered in areas with stronger social capital. Thus, as strong or severe as a formal public health rule is, social capital will encourage individuals to follow it. All else equal, this effect lowers the number of Covid-19 cases.

[[[Do we control for beliefs and normative values, e.g., red/blue states? That is one way to get at the difference between rows. Among red states do we see differences in social capital?!]]]

## III Data and Model

* Dependent Variable

  * Why do we focus on cases, not deaths?

* Independent Variable

## IV Results

Table 1 - Base Panel Regression Model of New cases as a function of stringency and stringency::social capital.

Table 2a - By Month
Table 2b - By every three months
Table 2c - eyeball date

Table 3 - IV

Table 4 - Controls and one-way fixed effects (time)

## V Conclusion
Notes:

* Social capital consistently influences infectious and preventative behaviors. 

* Informal rules like those found in social capital influence behavior and bolster formal rules.

* Policy implications are difficult to tease out given social capital is emergent. Social capital is a deeper notion than merely doing what an elected official tells us to do. Trusted advisers, business contacts, friends, and family comprise our social networks and capital.



While we can easily identify measures of social capital, e.g., the number of community centers or the percent of people who donate to charitable organizations, cultivating social capital---in the way it is meant in the literature---is not so simple. Perhaps for future outbreaks, we can discuss the role social capital might have and how to bolster that capital, but it seems unlikely that it will come from a centralized source.

* Future research should study the extent to which social capital erodes over time particularly because of formal rules that prevent interactions. That is, while we find social capital bolsters formal rules, those formal rules might extinguish the necessary parts of social capital given a longer duration or additional severity. This is especially once we recognize the cultural context within which social capital emerges, i.e., human interactions and the norms of proper behavior that develop. Formal rules that make such interactions more costly are likely to alter the costs and benefits people face such that existing social capital erodes (Carilli, Coyne, and Leeson, 2008) and alternative kinds of social capital will arise. Indeed, if formal rules prevent interactions that would have otherwise taken place, individuals realize less of a benefit from maintaining existing social ties. 

Alternatively, emphasis should be placed on providing quality information about Covid-19 transmission, as well as encouraging individuals to seek out relevant tests. Such measures encourage individuals to make decisions and *invest* in existing social ties according to their values.

* If two states have a stringency index of "6" - do they have the same policies? What are the potential differences? Same with social capital...

* Supporting Evidence:

  * Within states that have the worst social capital, are there pockets that have high social capital? 
  
  *Similarly, among the states that have really high social capital, are there pockets of really low social capital?